---
title: "Coding workshop: Week 10"
author: "your name here"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

```{r set-up}
# model packages
library(MASS) # have to read this in before tidyverse 
library(lme4)
library(glmmTMB) # ok if you don't have this - just comment it out

# diagnostics and model info
library(DHARMa)
library(MuMIn)
library(ggeffects)
library(lmtest)
library(broom)

# general usage
library(tidyverse)
library(here)
library(naniar)
library(skimr)
library(GGally)
library(flextable)

salamanders <- read_csv(here("data", "salamanders.csv"))
```

# Data info from `glmmTMB`:

`site`: name of a location where repeated samples were taken  
`mined`: factor indicating whether the site was affected by mountain top removal coal mining  
`cover`: amount of cover objects in the stream (scaled)  
`sample`: repeated sample  
`DOP`: Days since precipitation (scaled)  
`Wtemp`: water temperature (scaled)  
`DOY`: day of year (scaled)  
`spp`: abbreviated species name, possibly also life stage  
`count`: number of salamanders observed  

# Explore the data set:

## histogram of counts:

```{r histogram}
ggplot(data= salamanders, aes(x=count))+
  geom_histogram(bins=17)
```

## Missingness:

```{r missingness}
gg_miss_var(salamanders)
```

## Skim: 

```{r skim-dataset}
skim(salamanders)
```

## Pairs plot:  

```{r pairs-plot}
#| fig.width: 12
#| fig.height: 12
#| out.width: 90%
#| fig.align: center

salamanders %>% 
  #take out the site column
  select(!site) %>% 
  ggpairs()
```

# Question: How does salamander count vary with mined status, species, and stream cover?

## Build models

```{r models}
# linear model, we know this is wrong (because the data is not continuously distributed--this data is count/discrete: we are still doing this to show how it doesn't work)
salmad1 <- lm(count ~ cover + spp + mined, data= salamanders)

# generalized linear model with Poisson distribution (assumption that the mean of the assumes values is equal to the variance)
salmad2 <- glm(count ~ count + mined + spp, data = salamanders, family = "poisson")

# generalized linear model with negative binomial distribution (mean = mean, variance = mean divided by K or something (variance is greater than the mean)--> notes from this week)
salmad3 <- glm.nb(count ~ cover + mined + spp, data = salamanders)

# generalized linear model with Poisson distribution and random effect of site


# generalized linear model with negative binomial distribution and random effect of site

```

## Look at residuals

```{r model-residuals}
#| fig.height: 6
#| fig.width: 8
#| fig.align: center

# check diagnostics, resid v predict is synonomus to the resid v fitted 
simulateResiduals(salmad1, plot = TRUE) #this looks bad, overdispersed 
simulateResiduals(salmad2, plot = TRUE) #looks better than last plot, qq plot looks bad, over dispersed and outliers 
simulateResiduals(salmad3, plot = TRUE) #much better, residuals conform to uniform distribution, no sig problems detected ---> negative bionomial is likely the best fit 
```

## Which distribution to use?

```{r choosing-model}
model.sel(salmad1, salmad2, salmad3)
#minded and spp are categorical (+)
#cover is continuous (number)
#the weight-=1 indicates that is it 100% the best model, AICc also tells us this 
```

## Model summary

```{r summaries}
# model object
salmad3

# summary 
summary(salmad3)
#cover is not a sig predictors of salamander count 
#the amount of salamanders found on a mined mountain is less than an unmined mountain

# confidence intervals
confint(salmad3)

# adjusted R2
#variance in response explained by the variance in the predictors (r2m)
r.squaredGLMM(salmad3) #only need to look at the delta (43%-- faily low bc we are only using 3 predictors)
```

```{r}
# model object in table
salmad3 %>% 
  as_flextable()

#alternatively 
tidy(salmad3) %>% 
  flextable()
```


## Visualizing:

```{r visualization}
#| fig.width: 12
#| fig.height: 10
#| out.width: 90%
#| fig.align: center

plot(ggpredict(salmad3, terms= c("cover", "mined")), add.data = TRUE)
#point represent the raw data (there is a jitter applied to them)
#more salamanders on unmined mountaintops as opposed to mined mountaintops 

#included spp here because spp also is a predictor in cover in addition to mind
plot(ggpredict(salmad3, terms= c("cover", "mined", "spp")), add.data = TRUE)
```

